---
title: "Lab3"
author: "Blanco Romain, Carlin Cyril, Dong Stéphane"
date: "03/27/2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r pressure, echo=FALSE}
NAm2 <- read.table ("NAm2.txt", header = T)
cont<-function(x)
  {
    if (x %in% c("Canada"))
      cont<-"NorthAmerica"
    else if (x %in% c("Guatemala","Mexico","Panama","CostaRica"))
      cont<-"CentralAmerica"
    else
      cont<-"SouthAmerica"
    return (factor(cont))
  }
contID<-sapply(as.character(NAm2[,4]),FUN=cont)
```

### Question a

```{r, warning=FALSE, error=TRUE}
# Libraries needed
library("class")
library("MASS")

NAcont<-cbind(contID=contID,NAm2[,-(1:8)])
NAcont[,1]<-factor(NAcont[,1])
labels <-rep (1: 2, each=494/2); set = sample(labels, 494)
# Doesn't work because the variance-covariance matrix is not inversible
ldaModel <- lda(contID~., NAcont[1:1000,], subset=which(set==1))
```

Actually, R could not solve the matrix inverse because the within-class covariance matrix was singular thus the LDA doesn't work. That is why the lda function throw an error saying some variables appear to be constant within groups.

### Question b

```{r, warning=FALSE, echo=FALSE}
# We remove markers that have a null variance
withinvar<-apply(NAcont[(set==1),-1],FUN=
                   function(x){tapply(x,NAcont[(set==1),1],FUN=var)},MARGIN=2)
bool<-as.logical(apply(withinvar,FUN=function(x){prod(x!=0)},MARGIN=2))
NAcont2<-cbind(contID=contID,(NAm2[,-(1:8)])[,bool])
NAcont2[,1]<-factor(NAcont[,1])
cat(length(NAcont2), " markers on 5709 have not a null variance.")
```
It is slightly more than half of our dataset. This means this latter may not be suitable for the LDA method. 

```{r, warning=FALSE}
fit <- lda(contID~., NAcont2, subset = which(set==1))
summary(fit)
```

### Question c

```{r}
# Use of model to predict test set
prediction <- predict(fit, NAcont2[which(set==2),], type=c("class"))
summary(prediction)
```

### Question d

```{r}
# Confusion matrix
confMatrix <- table(NAcont2[which(set==2),1],prediction$class)
confMatrix
# Display barplot
mycol<-c("black","grey40","lightgrey")
barplot(t(confMatrix),col=mycol)
legend("topleft",paste(1:3),col=mycol,lwd=2,cex=2)
```


The barplot shows the succes rate of our prediction. As regards the North America, the confusion is big because the three colors are present. The reason could be the fact that the sample is small. As regards the Central America, there is a confusion between two continents the Central and the South America. Finally, the South America is well predicted by the model.

### Question e

```{r}
missclass<-function(x) {
  errorMean = 0;
  for(i in 1:length(x[1,]))
    for(j in 1:length((x[,1])))
      if(i != j)
        errorMean = errorMean + x[i,j]
  errorMean = errorMean/sum(x)
  print(errorMean)
}
missclass(confMatrix)
```
Around one third of the test set is missclassified. The model is clearly not accurate enough and needs to be improved. 

### Question f

As done in the previous Lab, we are going to operate a cross-validation method to improve the accuracy of our model and avoid overfitting results.

```{r,warning=FALSE}
# Operating a cross-validation with LDA method
l <- length(NAcont2$contID)
labels=rep(1:10,each=50, length.out = l)
set=sample(labels,l)
predictedID = matrix(nrow = l, ncol = 1)
for(i in 1:10) {
  fit <- lda(contID~., NAcont2, subset = which(set!=i))
  predictedID[which(set==i),1] <- (predict(fit,NAcont2[which(set==i),], type=c("class")))$class
}

# Confusion Matrix
confMatrix2 <- table(NAcont2[,1],predictedID)
confMatrix2

# Display barplot
mycol<-c("black","grey40","lightgrey")
barplot(t(confMatrix2),col=mycol)
legend("topleft",paste(1:3),col=mycol,lwd=2,cex=2)

# Rate of missclassified individuals
missclass(confMatrix2)
```
We can see we have improved our model for it is more accurate (about twice less), and the cross-validation method insure we don't have an overfitting model. The rate of missclassification can still be improved and we may consider another method than the LDA one.